Symbolic linearization of equations of motion of constrained multibody systems

### 2014-01-19 PetersonGedeHubbard2012.tex
Updated drawing, minor changes in text

\begin{abstract}
% Context
Many common dynamic mechanical systems have configuration or velocity constraints.
% Need  (what we have and what we want)
Analyses of such systems often require linearized forms of the motion
equations.
% Task
To address this issue, we developed a procedure for organizing the constraint
and motion equations and their subsequent linearization.  This procedure was
specifically developed for equations of motion generated by Kane's method, but
is compatible with any method as long as the following are true: only ordinary
differential equations are required to describe the system, and independent and
dependent states remain in the final equations (i.e. no DAE's, no substitution
of variables).
% Object of the document
Following a brief review of Kane's method and the structure of the equations it
generates, we present the procedure to symbolically linearize the nonlinear
motion equations of constrained multibody systems, and illustrate it with the
familiar example of the rolling disk.

@@

For a system system with constraints, Kane's dynamical equations\cite{Kane1985}
are first order ODE's (in the generalized speeds) which include both the
independent and dependent states (assuming the dependents states have not been
algebraically eliminated). The kinematical and dynamical differential equations
can be solved for the time derivatives of the generalized coordinates and
generalized speeds, respectively. Taken together, it is tempting to think that
this system of first order differential equations could be linearized in the
same fashion as an unconstrained system.  Unfortunately, this is not the case.
Linearizing these equations in the same manner as one would for an
unconstrained system (by directly computing the Jacobian) is incorrect and will
yield incorrect linearized equations (and subsequently incorrect eigenvalues,
in the case of a stability analysis). We illustrate this by means of a familiar
example, the rolling disk, formulated with a non-traditional set of coordinates
and generalized speeds which are not all independent. We provide a systematic
method to correctly generate the linear equations of motion for constrained
systems which have been found using Kane's method (or any other method which
results in equations with the same form). We apply this method to the rolling
disk example to illustrate that it yields eigenvalues that match previously
published results.

We consider three types of constraints: configuration, velocity, and
acceleration. Configuration constraints limit the location or orientation of
parts of the system, relative to the external world or other parts of the
system. Velocity constraints limit the speeds at which the configuration can
change, either from configuration constraints (through time differentiation) or
independent application. Velocity constraints most often appear in systems
where there is rolling without slip or there are closed kinematic loops. In the
context of this paper, acceleration constraints arise either from time
differentiated velocity constraints or similar kinematic considerations.

@@

% TODO:reiterate limitations of model
The goal of this paper is to definitively establish how the first time
derivative of all coordinates and speeds (independent \textit{and} dependent)
depend, to first order, upon a selection of independent coordinates and
independent speeds, for an arbitrary point of linearization. We demonstrate how
a reduced portion of these linearized equations can be used to perform standard
stability analysis (i.e., for control system design or linear stability
analysis). The formalism presented here is generic enough to cover most
examples of time-varying constrained multibody systems with arbitrary external
inputs and arbitrary specified quantities.
% TODO: electronic supplementary material with examples for other methods
% TODO: rewrite following sentences after e.s.m.
The procedure was created for systems whose equations of motion have been
derived with Kane's method. Although also applicable to dynamic system
equations formulated using other methods, it is restricted to systems which can
be completely described by a set of ODE's (i.e., DAE's needn't be solved).

>>>>

\begin{abstract}
% Context
Many common dynamic mechanical systems are subject to configuration or velocity
constraints.
% Need  (what we have and what we want)
Analyses of such systems often require linearized forms of the motion
equations.
% Task
To address this issue, we developed a procedure for organizing the constraint
and motion equations and their subsequent linearization.  This procedure was
specifically developed for equations of motion generated by Kane's method, but
is compatible with any method as long as the following are true: only ordinary
differential equations are required to describe the system, and independent and
dependent states remain in the final equations (i.e. no DAE's, dependent
variables are not substituted away).
% Object of the document
Following a brief review of Kane's method and the structure of the equations it
generates, we present the procedure to symbolically linearize the nonlinear
motion equations of constrained multibody systems, and illustrate it with an
example of the rolling disk.

@@

For a system with constraints, Kane's dynamical equations\cite{Kane1985} are
first order ODE's (in the generalized speeds) which include both the
independent and dependent states (assuming the dependent states have not been
algebraically eliminated). The kinematical and dynamical differential equations
can be solved for the time derivatives of the generalized coordinates and
generalized speeds, respectively. Taken together, it is tempting to think that
this system of first order differential equations could be linearized in the
same fashion as an unconstrained system.  Unfortunately, this is not the case.
Linearizing these equations in the same manner as one would for an
unconstrained system (by directly computing the Jacobian) is incorrect and will
yield incorrect linearized equations (and subsequently incorrect eigenvalues,
in the case of a stability analysis). We illustrate this by means of a familiar
example, the rolling disk, formulated with a non-traditional set of coordinates
and generalized speeds which are not all independent. We provide a systematic
method to correctly generate the linear equations of motion for constrained
systems which have been found using Kane's method (or any other method which
results in equations with the same form). We apply this method to the rolling
disk example to illustrate that it yields eigenvalues that match previously
published results.

We consider three levels of constraints: configuration, velocity, and
acceleration. Configuration (holonomic) constraints limit the location or
orientation of parts of the system, relative to the external world or other
parts of the system. Velocity (nonholonomic of time differentiated holonomic)
constraints limit the speeds at which the configuration can change.
Velocity constraints most often appear in systems where there is rolling
without slip or there are closed kinematic loops. For this procedure,
acceleration level constraints are assumed to be time differentiated velocity
constraints.

@@

% TODO:reiterate limitations of model
This paper presents linear, first order differential equations relating the
time derivatives of \textit{all} speeds and coordinates to the independent
speeds and coordinates, for any point of linearization.
A provided example demonstrates the necessity of these equations.
We demonstrate how a reduced portion of these linearized equations can be used
to perform standard stability analysis (i.e., for control system design or
linear stability analysis).
The formalism presented here is generic enough to cover most
examples of time-varying constrained multibody systems with arbitrary external
inputs and arbitrary specified quantities.
% TODO: electronic supplementary material with examples for other methods
% TODO: rewrite following sentences after e.s.m.
The procedure was created for systems whose equations of motion have been
derived with Kane's method. Although also applicable to dynamic system
equations formulated using other methods, it is restricted to systems which can
be completely described by a set of ODE's (i.e., DAE's needn't be solved).
